
---
title: "EDA Final"
author: "Jason Herbaugh"
date: "2/27/2021"
output:
  html_document:
    self_contained: false
    lib_dir: libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this case study we assume that our audience is the CEO and CFO of Budweiser (our client). They have hired us to answer 7 questions and beyond those general questions we will speculate / anticipate what may be of interest to them.


```{r}
options(scipen=999) #prevent scientific notation
#Load Libraries needed for analysis
library(dplyr)
library(tidyverse)
library(visdat)
library(GGally)
library(usmap)
library(mice)
library(VIM)
library(plotly)
library(ggpubr)
library(caret)
library(e1071)
library(class)
library(maps)
library(mapproj)
library(stringr)
library(ggplot2)
library(ggthemes)
library(table1)
library(ggcorrplot)
library(mlbench)
library(Boruta)
library(e1071)
library(MASS)
library(naivebayes)
library(mboost)
library(kernlab)
library(randomForest)
library(doParallel)
library(HH)
```

Load Theme for plots

```{r}
theme_set(theme_light())
theme_update(axis.title = element_text()) #the default for fivethirtyeight is to not show axis labels, this removes that default so we can choose to specify and display axis titles
theme_update(plot.title = element_text(hjust = 0.5)) # changing default to center all titles
```

Load initial data sets provided

```{r}
talent_raw = read.csv("https://raw.githubusercontent.com/jherbaugh/CaseStudy2DDS/main/CaseStudy2-data.csv", header = TRUE)
talent_attrition_test = read.csv("https://raw.githubusercontent.com/jherbaugh/CaseStudy2DDS/main/CaseStudy2CompSet%20No%20Attrition.csv", header = TRUE)
talent_salary_test = read.csv("https://raw.githubusercontent.com/jherbaugh/CaseStudy2DDS/main/CaseStudy2CompSet%20No%20Salary.csv", header = TRUE)


#Read in Dimension of the Dataset
dim(talent_raw) #870 observations x 36 columns


#Summary Statistics and Glimpse into DataTypes and first observations
glimpse(talent_raw)
summary(talent_raw)
```

Convert character to factors, drop columns with no variance
```{r}
talent_factor <- talent_raw %>%
  mutate_if(is.character, as.factor) %>%
  dplyr::select(Attrition, everything(), -Over18, -EmployeeCount, -StandardHours)

```

Investigate NA values to determine what needs resolution (Nothing is missing)

```{r}
#Verify no missing observations
md.pattern(talent_factor)

#Summary Statistics and Glimpse into DataTypes and first observations for Training
glimpse(talent_factor)
summary(talent_factor)
```
Summary Statistics on additional Data Frames

```{r}
#Summary Statistics and Glimpse into DataTypes and first observations for Test sets
summary(talent_attrition_test)
summary(talent_salary_test)

```
Attrition Properties in Dataset

```{r}
prop.table(table(talent_factor$Attrition))

```

Create data frame of continuous variables and summarize means values for attrition factor. Creation Correlation Matrix for continuous variable.

```{r}

talent_continuous <- talent_factor[, c("YearsWithCurrManager", "YearsSinceLastPromotion", "YearsInCurrentRole", "YearsAtCompany", "PerformanceRating",
                                       "TotalWorkingYears", "PercentSalaryHike", "MonthlyIncome", "MonthlyRate", "HourlyRate", "DailyRate", "WorkLifeBalance", "OverTime",
                                       "DistanceFromHome", "Age", "Attrition", "JobSatisfaction", "JobLevel", "RelationshipSatisfaction", "EnvironmentSatisfaction")]


talent_continuous$Attrition <- as.numeric(talent_continuous$Attrition) - 1
talent_continuous$OverTime <- as.numeric(talent_continuous$OverTime) - 1

employees_left <- talent_continuous %>% filter(talent_continuous$Attrition == 1)
employees_stayed <- talent_continuous %>% filter(talent_continuous$Attrition == 0)
mean_employees_left <- lapply(employees_left, mean)
mean_employees_stayed <- lapply(employees_stayed, mean)

mean_employees <- merge(mean_employees_left, mean_employees_stayed, all=TRUE)
mean_employees

# Age - Left Company (Y/N)
# - Yes = 33.78
# - No = 37.41
#
# Monthly Income
# - Yes = $4,764
# - No = $6,702
#
# Job Satisfaction | Environment Satisfaction | Relationship Satisfaction
# - Yes = 2.44       2.51                       2.61
# - No = 2.76        2.73                       2.72
#
# Years with Current Manager | Years Since Last Promo | Years in Current Role
# - Yes = 2.94                 2.13                     2.90
# - No = 4.369863              2.17                     4.54
#
#  Total Working Years
# - Yes = 8.19
# - No = 11.60

# Performance Rating | Percent Salary Hike
# - Yes = 3.16         15.33
# - No = 3.15          15.17  
#
# Work Life               | Dist from Home
# - Yes = 2.63              10.96
# - No = 2.80               9.02

```

Correlation Matrix for data

```{r}

correlations <- round(cor(talent_continuous), 1)

# method = "circle"
ggcorrplot(correlations, hc.order = TRUE, type = "lower",
           lab = TRUE)

# During review of the correlation matrix, we can see high instances of correlation in the variables below.
# They will be removed in future analysis.
#
# High collinearity:
# - Performance Rating with Percent Salary Hike
# - Monthly Income with Job Level
# - Age and Total Working years
# - Years at Current Role with Years at Company
# - Years at Company with Years with Current Manager
#
# Variables to drop :
# - Percent Salary Hike
# - Job Level
# - Total Working Years
# - Years at Company

```

Additional insights from data
```{r}
talent_factor$agegroup <- with(talent_factor, ifelse(Age < 25, "18-24", ifelse(Age < 34, "25-34", ifelse(Age < 44, "35-44", ifelse(Age < 52, "45 - 52", "Over53")))))
talent_attrition_test$agegroup <- with(talent_attrition_test, ifelse(Age < 25, "18-24", ifelse(Age < 34, "25-34", ifelse(Age < 44, "35-44", ifelse(Age < 52, "45 - 52", "Over53")))))
talent_salary_test$agegroup <- with(talent_salary_test, ifelse(Age < 25, "18-24", ifelse(Age < 34, "25-34", ifelse(Age < 44, "35-44", ifelse(Age < 52, "45 - 52", "Over53")))))
age_count <- table(talent_factor$agegroup)
barplot(age_count, main = "Age Count")


### Mean Salary for Job Title/ Education and Salary

talent_factor_jobtitle_mean <- talent_factor %>% dplyr::group_by(JobRole) %>% dplyr::summarise(MedSalary = median(MonthlyIncome, na.rm=TRUE),Meansatis= mean(JobSatisfaction, na.rm=TRUE))
#Subsetting the job satisfaction data into top/bottoms
mean_jobsatisfaction <- talent_factor_jobtitle_mean %>% arrange(desc(Meansatis))
#Subsetting the salary data into top/bottoms
mean_salary <- talent_factor_jobtitle_mean %>% arrange(desc(MedSalary))
#Mean Job Satisfaction by Job Role
ggplot(mean_jobsatisfaction, aes(x=reorder(JobRole,-Meansatis),y=Meansatis,fill=JobRole)) + geom_bar(stat = "identity")  + theme(plot.title = element_text(hjust=0.1), axis.text.x = element_text(hjust = 0.9, angle = 65)) + labs(x="Job Role", y="Mean Job Satisfaction", title = "Job Satisfaction by Title")
#Median Income by Job Role
ggplot(mean_salary, aes(x=reorder(JobRole,-MedSalary),y=MedSalary,fill=JobRole)) + geom_bar(stat = "identity")  + theme(plot.title = element_text(hjust=0.1), axis.text.x = element_text(hjust = 0.9, angle = 65)) + labs(x="Job Role", y="Median Monthly Salary", title = "Median Monthly Salary by Title")


#Attrition plotting
Attritionplot <- ggplot(talent_factor, aes(x=DailyRate, y=MonthlyIncome), title ='Attrition') + geom_point()
Attritionplot + facet_wrap(~Attrition)
Satisfactionplot <- ggplot(talent_factor, aes(x=DailyRate, y=MonthlyIncome), title ='Attrition') + geom_point()
Attritionplot + facet_wrap(~JobSatisfaction)
Attritionsat <- ggplot(talent_factor, aes(x=JobSatisfaction, y=MonthlyIncome), title ='Attrition') + geom_point()
Attritionsat + facet_wrap(~Attrition)
ggplot(talent_factor, aes(x=JobRole, fill=Attrition)) + geom_bar() +  labs(x = "Job Role", y = "Number of Employees", title = "Attrition Job Role")


# Distance from home grouping
talent_factor$DistanceGrouping <- with(talent_factor, ifelse(DistanceFromHome > 25, "25+ Miles", ifelse(DistanceFromHome > 18, "19 - 25 Miles", ifelse(DistanceFromHome > 10, "11 - 18 Miles", ifelse(DistanceFromHome >5, "6 - 10 Miles", "Less than 6  Miles")))))
talent_attrition_test$DistanceGrouping <- with(talent_attrition_test, ifelse(DistanceFromHome > 25, "25+ Miles", ifelse(DistanceFromHome > 18, "19 - 25 Miles", ifelse(DistanceFromHome > 10, "11 - 18 Miles", ifelse(DistanceFromHome >5, "6 - 10 Miles", "Less than 6  Miles")))))
talent_salary_test$DistanceGrouping <- with(talent_salary_test, ifelse(DistanceFromHome > 25, "25+ Miles", ifelse(DistanceFromHome > 18, "19 - 25 Miles", ifelse(DistanceFromHome > 10, "11 - 18 Miles", ifelse(DistanceFromHome >5, "6 - 10 Miles", "Less than 6  Miles")))))
ggplot(talent_factor, aes(x=reorder(DistanceGrouping, DistanceFromHome), fill=Attrition)) + geom_bar()  + labs(x = "Distance From Home Group", y = "Number of Employees", title = "Attrition and Distance from Home")

```

GGPairs broken down by various Talent Factor classes

```{r}
##GGPairs Correlation matrix broken down into different themed groups
##Observations based upon the output

#Employee measurements
#People that left were younger
#People that left were less educated
talent_employee <- talent_factor %>% dplyr::select(Attrition, Age, Education, Gender, MaritalStatus)
ggpairs(talent_employee, aes(colour = Attrition))


#Employer based measurements
#Pople who left had less time in role , at company, job level, enviornment satisfication
talent_company <- talent_factor %>% dplyr::select(Attrition, EnvironmentSatisfaction, JobInvolvement, YearsInCurrentRole, YearsAtCompany , JobLevel)
ggpairs(talent_company, aes(colour = Attrition))


#EmployeeWorkLifeBalance
#Maybe distance from home
talent_wellbeing <- talent_factor %>% dplyr::select(Attrition, BusinessTravel, DistanceFromHome, WorkLifeBalance)
ggpairs(talent_wellbeing, aes(colour = Attrition))

#Income Based Grouping
#Not Seeing strong correlation with attrition in relation to salary. Drop OverTime, StockOptionLevel no real info. Add SalaryHike
talent_income <- talent_factor %>% dplyr::select(Attrition, DailyRate, HourlyRate, MonthlyRate, MonthlyIncome, PercentSalaryHike )
ggpairs(talent_income, aes(colour = Attrition))

```

Scatterplots of additional data points

```{r}
# People who left salary left made less over time
ggscatter(
  talent_factor, x = "TotalWorkingYears", y = "MonthlyIncome",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()

# People who left made less to start but slalary increased same
ggscatter(
  talent_factor, x = "YearsAtCompany", y = "MonthlyIncome",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# People who left had fewer total working year and years at company
ggscatter(
  talent_factor, x = "TotalWorkingYears", y = "YearsAtCompany",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()

# People who left had fewer years with current manager / current role (to start with)
ggscatter(
  talent_factor, x = "YearsWithCurrManager", y = "TotalWorkingYears",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# People who left had fewer years in current role (to start with)
ggscatter(
  talent_factor, x = "YearsInCurrentRole", y = "TotalWorkingYears",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# Attrition -> People who left at 1/2 as much time at current company (to start)
ggscatter(
  talent_factor, x = "YearsWithCurrManager", y = "YearsAtCompany",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# Attrition -> People who left at 1/2 as much time at current company (to start)
ggscatter(
  talent_factor, x = "YearsInCurrentRole", y = "YearsAtCompany",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# Attrition -> People with left had 1/2 the time with current manager since last promotion (same as below, can remove one)
ggscatter(
  talent_factor, x = "YearsSinceLastPromotion", y = "YearsInCurrentRole",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# Attrition -> People with left had 1/2 the time with current manager since last promotion
ggscatter(
  talent_factor, x = "YearsSinceLastPromotion", y = "YearsWithCurrManager",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


# Attrition -> High correlation with same manager and number of years in current role
ggscatter(
  talent_factor, x = "YearsInCurrentRole", y = "YearsWithCurrManager",
  color = "Attrition", palette = "jco",
  add = "reg.line"
) +
  facet_wrap(~Attrition) +
  stat_cor(label.y = 5.4) +
  stat_regline_equation()


```

Modeling our Analysis
Global Values

```{r}
#This will be the dataset we use to model our Attrition and Salary Prediction
set.seed(453)

### Removed high correlation variables from set
splitPerc = .70


talent_factor_model <-  talent_factor %>% dplyr::select(everything(), -YearsAtCompany, -TotalWorkingYears, -JobLevel, -PercentSalaryHike )
trainInd <- sample(1:dim(talent_factor_model)[1],round(splitPerc * dim(talent_factor_model)[1]))

talent_factor_model_Train = talent_factor_model[trainInd,]
talent_factor_model_Test = talent_factor_model[-trainInd,]
```

Attrition Classification - Linear Regression

```{r}
#Linear Model = need more on this
talent_attrition_lm <- lm(Attrition ~., data=talent_factor_model_Train)
talent_attrition_lm
```

Naive Bayes

```{r}
NB_model = naiveBayes(talent_factor_model_Train[,-1],as.factor(talent_factor_model_Train$Attrition),laplace = 1)
table(predict(NB_model,talent_factor_model_Test[,-1]),as.factor(talent_factor_model_Test$Attrition))
CM = confusionMatrix(table(predict(NB_model,talent_factor_model_Test[,-1]),as.factor(talent_factor_model_Test$Attrition)))

CM
```

Random Forest

```{r}
RF_talent_factor_model_train <- train(Attrition ~., talent_factor_model_Train, method = 'rf', trControl = trainControl(method='repeatedcv'), importance = T)

predTrain <- predict(RF_talent_factor_model_train, talent_factor_model_Test)
confusionMatrix(table(talent_factor_model_Test$Attrition, predTrain))
cMatrixRF <- table(predTrain, talent_factor_model_Test$Attrition)
plot(cMatrixRF, col="blue", ylab="Actual", xlab="Predicted", main='Random Forest Confusion Matrix')



```

SVM Model

```{r}


svm_talent_factor_model_train <- train(Attrition ~., talent_factor_model_Train, method = 'svmRadial', trControl = trainControl(method='repeatedcv'), importance=T)
predTrainsvm <- predict(svm_talent_factor_model_train, talent_factor_model_Test)
confusionMatrix(table(talent_factor_model_Test$Attrition, predTrainsvm))
cMatrixsvm <- table(predTrainsvm, talent_factor_model_Test$Attrition)
plot(cMatrixsvm, col="blue", ylab="Actual", xlab="Predicted", main='Support Vector Machine Confusion Matrix')

```

Feature Engineering for importance variable related to attrition
```{r}
### Feature Importance for Predictive ability
# ensure results are repeatable
# load the library


# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)

# train the model
results <- rfe(talent_factor[,2:33], talent_factor[,1], sizes=c(2:33), rfeControl=control)

# summarize the results
print(results)

# list the chosen features
predictors(results)

# plot the results
plot(results, type=c("g", "o"))


#Boruta Feature Engineering
boruta.talent_train <- Boruta(Attrition~., data = talent_factor, doTrace = 2)
print(boruta.talent_train)

#take a call on tentative features
boruta.talent <- TentativeRoughFix(boruta.talent_train)
print(boruta.talent)

getSelectedAttributes(boruta.talent_train, withTentative = F)

plot(boruta.talent, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta.talent$ImpHistory),function(i)
  boruta.talent$ImpHistory[is.finite(boruta.talent$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.talent$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
     at = 1:ncol(boruta.talent$ImpHistory), cex.axis = 0.7)


```

Salary Prediction

```{r}
salaryfitincome <- lm(MonthlyIncome ~ ., data = talent_factor_model_Train)
summary(salaryfitincome)
vif(salaryfitincome)


talent_factor_model_salary <-  talent_factor %>% dplyr::select(everything(), -Age, -DistanceFromHome, -Department )
trainInd <- sample(1:dim(talent_factor_model_salary)[1],round(splitPerc * dim(talent_factor_model_salary)[1]))

talent_factor_model_salary_Train = talent_factor_model_salary[trainInd,]
talent_factor_model_salary_Test = talent_factor_model_salary[-trainInd,]

##Interaction rates of all the variables to see collinearity
## Variables where we are seeing high collinearity within categorical variables, we will leave in. 
##Significant p-value of <0.0001.

#Variables with High collinearity

# Age/Age Group
# Distance Grouping
# Department



```

Salary Prediction - LinearRegression

```{r}
#Job role coefficients are going to be the biggest predictors of monthly income. followed by business travel and performance rating.
salary_predict_lm <- lm(MonthlyIncome ~., data= talent_factor_model_salary_Train)
summary(salary_predict_lm)
vif(salary_predict_lm)


```

Salary Prediction - KNN
```{r}
#Our Lowest RMSE came with a K nearest neighbors of 7. 
#This prediction model gives us an RMSE of $2,067 on monthly income, which falls below our $3,000 evaluation criteria.
traind <- trainControl(method = 'repeatedcv', number = 10, repeats = 3)
knn_fit <- train(MonthlyIncome ~ ., data = talent_factor_model_salary_Train, method = 'knn', trControl = traind, preProcess = c('center', 'scale'), tuneLength = 10)
knn_fit
plot(knn_fit)
predregtrainknn <- predict(knn_fit, talent_factor_model_salary_Test)


```

Salary Prediction - Random Forest

```{r}

# Our optimal value for RMSE is $1037. However, we may be oveerfitting.

rf_fit <- train(MonthlyIncome ~ ., data = talent_factor_model_salary_Train, method = 'rf', trControl = traind, preProcess = c('center', 'scale'), tuneLength = 10, importance=T)
rf_fit
plot(rf_fit)
predregtrainrf <- predict(rf_fit, talent_factor_model_salary_Test)



```


Salary Prediction - SVM 
```{r}
### SVM METHOD RMSE IS 1366 and Rsquared is .91

svm_fit <- train(MonthlyIncome ~ ., data = talent_factor_model_salary_Train, method = 'svmRadial', trControl = traind, preProcess = c('center', 'scale'), tuneLength = 10, importance=T)
svm_fit
plot(svm_fit)
predregtrainsvm <- predict(svm_fit, talent_factor_model_salary_Test)


```

Feature Engineering for importance variable related to salary

```{r}
### Variable importance of our regressor
rfimportreg <- varImp(rf_fit, scale = FALSE)

plot(rfimportreg)


# Job Role, Age, and then Years with Current Manager are the top 3 important variables.. 
# Job title is typically a heavier predictor in salary and the older you are, the more senior you typically are, which affects your Salary. 
#We also have Years with Current Manager closely followed by Years in Current Role, which would make sense as it can tie back to a users job performance and raises #once they have plenty of experience working the same job with the same boss.



```

Run attrition classification on Test Data and output to csv
```{r}
#Execute Classification model against test set and write out to csv
talent_attrition_predict<- predict(RF_talent_factor_model_train, talent_attrition_test)
talent_attrition_merge <- cbind(talent_attrition_test$ID, talent_attrition_predict)
talent_attrition_class <- as.data.frame(talent_attrition_merge)
talent_attrition_class$Attrition <- with(talent_attrition_class, ifelse(talent_attrition_predict == 1, "No", "Yes"))
talent_attrition_final <- talent_attrition_class[,!(names(talent_attrition_class) %in% c("talent_attrition_predict"))]
#outputting data
write.csv(talent_attrition_final,"JasonHerbaugh_DDSCaseStudy2_PredictAttrition.csv",row.names = FALSE)

```

Run salary prediction on Test Data

```{r}
#run prediction on rf classification model using our test set
talent_salary_test$Attrition <- with(talent_salary_test, ifelse(Attrition == 1, "Yes", "No"))
talent_salary_regression <- predict(rf_fit, talent_salary_test)
talent_salary_merge <- cbind(talent_salary_test$ID, talent_salary_regression)
talent_salary_pred <- as.data.frame(talent_salary_merge)
names(talent_salary_pred)<- c("ID", "MonthlySalary")
#outputting data
write.csv(talent_salary_pred,"JasonHerbaugh_DDSCaseStudy2_PredictSalary.csv",row.names = FALSE)
```


Session Information

```{r}
sessionInfo()
```

